---
title: "Visualization of Correlation within Spatial Experimental Plots"
subtitle: "Graphics Group"
author: "Alison Kleffner"
date: "Department of Mathematics, Creighton University"
output:
  xaringan::moon_reader:
    seal: false
    includes:
      after_body:
        "js-addins.html"
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: ["default", "metropolis-fonts", "metropolis" ,"css/modal.css", "css/sizeformat.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightlines: true
      countIncrementalSlides: true
---
class:title-slide-custom

```{r, child = "style.Rmd"}
```


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
# Packages
library(emoji)
library(purrr)
library(tidyverse)
library(gridExtra)
library(nullabor)
library(scales)
library(knitr)
library(kableExtra)
library(RefManageR)
library(fontawesome)
library(shiny)
# download_fontawesome()

# References
bib <- ReadBib("bib/thesis.bib", check = FALSE)
ui <- "- "

# R markdown options
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      cache = TRUE,
                      dpi = 300)
options(htmltools.dir.version = FALSE)
options(knitr.kable.NA = '')
```

```{r, include = F, eval = T, cache = T}
clean_file_name <- function(x) {
  basename(x) %>% str_remove("\\..*?$") %>% str_remove_all("[^[A-z0-9_]]")
}
img_modal <- function(src, alt = "", id = clean_file_name(src), other = "") {
  
  other_arg <- paste0("'", as.character(other), "'") %>%
    paste(names(other), ., sep = "=") %>%
    paste(collapse = " ")
  
  js <- glue::glue("<script>
        /* Get the modal*/
          var modal{id} = document.getElementById('modal{id}');
        /* Get the image and insert it inside the modal - use its 'alt' text as a caption*/
          var img{id} = document.getElementById('img{id}');
          var modalImg{id} = document.getElementById('imgmodal{id}');
          var captionText{id} = document.getElementById('caption{id}');
          img{id}.onclick = function(){{
            modal{id}.style.display = 'block';
            modalImg{id}.src = this.src;
            captionText{id}.innerHTML = this.alt;
          }}
          /* When the user clicks on the modalImg, close it*/
          modalImg{id}.onclick = function() {{
            modal{id}.style.display = 'none';
          }}
</script>")
  
  html <- glue::glue(
     " <!-- Trigger the Modal -->
<img id='img{id}' src='{src}' alt='{alt}' {other_arg}>
<!-- The Modal -->
<div id='modal{id}' class='modal'>
  <!-- Modal Content (The Image) -->
  <img class='modal-content' id='imgmodal{id}'>
  <!-- Modal Caption (Image Text) -->
  <div id='caption{id}' class='modal-caption'></div>
</div>
"
  )
  write(js, file = "js-addins.html", append = T)
  return(html)
}
# Clean the file out at the start of the compilation
write("", file = "js-addins.html")
```

<br><br><br>
## Visualization of Correlation within Spatial Experimental Plots
#### Alison Kleffner
#### Department of Mathematics, Creighton University
##### `r fa("envelope", fill = "black")` [alisonkleffner@creighton.edu](email.address@creighton.edu)
##### `r fa("home", fill = "black")` [https://alison-kleffner.netlify.app/](https://alison-kleffner.netlify.app/)
##### `r fa("github", fill = "black")` [alisonkleffner](https://github.com/alisonkleffner)
<br><br>
.medium[*Slides: https://alisonkleffner.github.io/presentations-creighton/graphics-group/S24/index.html#1*]

---
class:primary
# Outline

`r fa_i("chart-bar")` Background

`r fa_i("list")` Current spatial visualizations for two variable comparisons 

`r fa_i("list")` Redesigning these spatial visualizations

---
class:primary
# Spatial Data

Spatial data relates to a geographical area or location.

- Often, more than one attribute is measured at a location.

**Focus**: However, due to the variables occupying the same space, multivariate spatial data is complex

- The spatial information provides additional context, so should be included
- Difficulties in visualization due to issues like clutter `r Citep(bib[[c("he-mult-2019")]])`

**Example**: Visualization of crop input application versus crop yield

---
class:primary
# Background

- With a projected increase in future crop demand, researchers are conducting studies on crop input application to increase yield, focusing on sustainability `r Citep(bib[[c("tilman_sustainalbe_2011")]])`

- Crop Input Example: Nitrogen Fertilizer
  + Nitrogen is an essential component of food production as allows plants to photosynthesize efficiently `r Citep(bib[[c("MAHESWARI2017175")]])`
  + Nearly half of the nitrogen fertilizer supplied to the field is not used by the crops `r Citep(bib[[c("billen_nitrogen_2013")]])`
  + This excess nitrogen can be harmful

--

**Data Collected:**

- Trial Design
- As-applied
- Yield
- Location of measurements


---
class:primary
# Comparative Graphics

`r Citet(bib[[c("gleicher2011")]])` surveyed journal articles and grouped comparative layouts into: superposition, juxtaposition, explicit encoding, or a hybrid of
two of the previous arrangements.

.center[
```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/comparative-graphs.png", alt = "A. Juxtaposition; B. Superposition; C. Explicit Encoding (Gleicher et al 2011)",other=list(width="25%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```
]

A). Juxtaposition: side-by-side
- Benefits: less issues with visual clutter and easier to create
- Drawbacks: comparative burden is placed on the user

B). Superposition: multiple objects in same coordinate system

- Benefits: Easier to compare as users can use perception rather than memory
- Drawbacks: clutter

C). Explicit Encoding: Directly encode relationship

- Benefits: Reader directly shown desired relationship
- Drawbacks: Lose data context

---
class:primary
# Current Example: Superposition

.center[

```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/old_map.jpg", alt = "Original Superimposed Plots.",other=list(width="25%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```

]

--

Issues with this Graph:

- Multiple dots on top of one another
  + Obscures true number of dots, harder to find patterns
  + Visual cues, like color, becomes partially obstructed, reducing search efficiency `r Citep(bib[[c("BRAVO2004a", "BRAVO2004b")]])`
  + Overburdens human perception, causing errors in performing tasks `r Citep(bib[[c("huang2009")]])`

---
class:primary
# Current Example: Juxtaposition

.center[

```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/juxtaposed-ex.jpeg", alt = "Orignial Juxtaposed Example.",other=list(width="40%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```

]

--

Issues with this Graph:

- Most of the comparative burden placed on users' memory
  + A mental image is relied on for comparison in these scenarios, as the user moves their eyes between images (shifting focus) `r Citep(bib[[c("visual-tominski-2012")]])`.
  + The plot contents may not be accurately formed in working memory, leading to potential errors when deriving patterns `r Citep(bib[[c("lyi2021")]])`
  + Lack of visual cues for locations
  
---
class:primary
# In Literature

.center[
```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/maxwell.png", alt = "Maxwell et al (2018)", other=list(width="40%"))
i2 <- img_modal(src = "images/trevisan.png",alt = "Trevisan et al. (2021)", other=list(width="40%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i2, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9],
  str_split(i2, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```
]

--

Color Schemes!


---
class:primary
# Redesign: Color Blending

.center[
```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/Attempt2.png", alt = "Color Blending with Original Yield Polygons", other=list(width="35%"))
i2 <- img_modal(src = "images/Attempt2-scale.png", alt = "Color Blending with Combined Yield Polygons.", other=list(width="30%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i2, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9],
  str_split(i2, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```
]


- Started with superposition: useful when spatial location is an essential element for deriving this information `r Citep(bib[[c("wang_comp18")]])`
- Suggestion: color blending `r Citep(bib[[c("macdonald_1999", "SILVA2011320")]])`
    + Though others claimed this may make comparisons more difficult `r Citep(bib[[c("visual-tominski-2012")]])`

---
class:primary
# Redesign: Bivariate Color Map

.center[

```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/biv-plot-scale.png", alt = "Bivariate Color Map Example.",other=list(width="20%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```

]

Useful when want to emphasize the relationship between two variables `r Citep(bib[[c("elmer2013symbol")]])`

- `r Citet(bib[[c("Leonowicz2003RESEARCHOT")]])` suggested a 2x2 or 3x3 color scheme
- `r Citet(bib[[c("wainer-1980")]])` suggests people have difficulty reading these maps correctly
  - Controversial color scheme, with 16 colors

**Focus**: diagonal
- Diagonal: grayscale color scheme.
- Upper left and lower right: complementary color scheme `r Citep(bib[[c("strode-2020")]])`


---
class:primary
# Redesign: Correlation Plot

.center[
```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/corr-plot2.png", alt = "Correlation Calculated within Original Trial Design Experimental Plots.", other=list(width="40%"))
i2 <- img_modal(src = "images/corr-plot2b-scale.png", alt = "Correlation Calculated within Larger Experimental Plots", other=list(width="30%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i2, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9],
  str_split(i2, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```
]

**Focus**: Wanted to give a more direct statement of correlation while maintaining some spatial orientation

- Bivariate color scale
- Maintain some spatial information: correlations may be impacted by field location

---
class:primary
# Redesign: Correlation Plot Interactive

.center[

```{r results='asis', echo = F, include = T, cache = T, eval = TRUE}
i1 <- img_modal(src = "images/interactive-new.jpg", alt = "Static version of Interactive Plot",other=list(width="50%"))

c(str_split(i1, "\\n", simplify = T)[1:2],
  str_split(i1, "\\n", simplify = T)[3:9]
  ) %>% paste(collapse = "\n") %>% cat()
```

]

[Link](https://alisonkleffner.github.io/yield-map-redesign/interactive-new.html)

A standard practice to overcome weakness of decontextualization is utilizing a hybrid comparative layout `r Citep(bib[[c("lyi2021")]])`
  - Juxtaposed scatterplot to the correlation plot (combining the layouts of juxtaposition and explicit encoding).
  - Used [ggiraph](https://davidgohel.github.io/ggiraph/)
  
---
class:primary
# Future Work

**Next Step**: Obtain Feedback from those using the plots

- Eventually do some testing between the layouts to see which farmers are reading more accurately.


[GitHub Link](https://github.com/alisonkleffner/yield-map-redesign)



---
class:primary
# References
<font size="1.5">
```{r, print_refs1, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
print(bib[[c("he-mult-2019", "tilman_sustainalbe_2011", "MAHESWARI2017175", "billen_nitrogen_2013", "gleicher2011", "BRAVO2004a", "BRAVO2004b", "huang2009", "visual-tominski-2012", "lyi2021", "wang_comp18", "macdonald_1999", "SILVA2011320", "elmer2013symbol", "wainer-1980", "Leonowicz2003RESEARCHOT", "strode-2020")]], 
      .opts = list(check.entries = FALSE, style = "html", bib.style = "authoryear")
      )
```
</font>

---
class:inverse
<br>
<br>
<br>
.center[
# Questions?
<br>
<br>
`r fa("envelope", fill = "white")` **alisonkleffner@creighton.edu**
`r fa("github", fill = "white")` **alisonkleffner**
]
